---
output:
  word_document: default
  html_document: default
---
```{r}
library(tidyverse) 
library(GGally)
library(MASS)
library(leaps)
```

```{r}
bike = read_csv("hour.csv")
```

```{r}
bike = bike %>% mutate(season = as_factor(as.character(season))) %>%
mutate(season = fct_recode(season,
"Spring" = "1",
"Summer" = "2",
"Fall" = "3",
"Winter" = "4"))
```

```{r}
bike = bike %>% mutate(yr = as_factor(as.character(yr))) %>%
  mutate(mnth = as_factor(as.character(mnth))) %>%
  mutate(hr = as_factor(as.character(hr)))
```

```{r}
bike = bike %>% mutate(holiday = as_factor(as.character(holiday))) %>%
mutate(holiday = fct_recode(holiday,"NotHoliday" = "0","Holiday" = "1")) 
```

```{r}
bike = bike %>% mutate(workingday = as_factor(as.character(workingday))) %>%
mutate(workingday = fct_recode(workingday,"NotWorkingDay" = "0","WorkingDay" = "1")) 
```

```{r}
bike = bike %>% mutate(weathersit = as_factor(as.character(weathersit))) %>%
mutate(weathersit = fct_recode(weathersit,"NoPrecip" = "1","Misty" = "2", "LightPrecip" = "3", "HeavyPrecip" = "4")) 
```

```{r}
bike = bike %>% mutate(weekday = as_factor(as.character(weekday))) %>%
mutate(weekday = fct_recode(weekday,"Monday" = "1","Tuesday" = "2", "Wednesday" = "3", "Thursday" = "4", "Friday" = "5", "Saturday" = "6", "Sunday" = "0")) 
```

Why do we convert “yr”, “mnth”, and “hr” into factors? Why not just leave them as numbers?
//To make it easier to run and implement in our future analysis. 


```{r}
ggpairs(bike, columns = c("temp", "atemp","hum", "windspeed", "count"))
```

Which of the quantitative variables appears to be best correlated with “count” (ignore the “registered”
and “casual” variable as the sum of these two variables equals “count”)?
// Temperature

```{r}
ggplot(bike,aes(x=hr,y=count)) + geom_boxplot()
ggplot(bike,aes(x=yr,y=count)) + geom_boxplot()
ggplot(bike,aes(x=mnth,y=count)) + geom_boxplot()
ggplot(bike,aes(x=holiday,y=count)) + geom_boxplot()
ggplot(bike,aes(x=weekday,y=count)) + geom_boxplot()
ggplot(bike,aes(x=workingday,y=count)) + geom_boxplot()
ggplot(bike,aes(x=weathersit,y=count)) + geom_boxplot()
```

Which variables appear to affect "count"? 

// yr = With only 2 different responses, is difficult to spot a trend, even though we can see an increment in the mean    from 2011 to 2012
   mnth = Month shows a fairly easy to spot trend to encrease count during the warmer months of the year.
   holiday = Is difficult to say that holiday affect count since we only have to differen responses, but according to    the plot, the mean number of count is greater in Not Holidays, as well as the deviation. 
   Weekday = We can see a slight trend between weekday and count, the weekend being the days with greater count.
   weathersit = we can see atrend where count decreases as the level of precipitation increases. 
   
```{r}
bike2 = bike %>% dplyr::select("count", "season", "yr", "mnth", "hr", "holiday", "weekday", "workingday", "weathersit", "temp", "atemp", "hum", "windspeed")
```

```{R}
allmod = lm(count ~., bike2)
summary(allmod)

emptymod = lm(count ~1, bike2)
summary(emptymod)
```

```{r}
forwardmod = stepAIC(emptymod, direction = "forward", scope=list(upper=allmod,lower=emptymod),
                      trace = TRUE)
summary(forwardmod)
```

What variables are included in your forward model? Comment on the quality of the model. Does the model
match our intuition/common sense?
// hr,atemp, yr, weathersit, season, mnth, hum, weekday, holiday, windspeed, temp. It is a fairly good model with an Adjusted R=squared of 0.6854 and good p value. For the quantitative variables the model makes sense, as for the categorical or dummy variables like year, it does not make sense, but the model still considered them imporant enough to include them.

```{r}
backmod = stepAIC(allmod, direction = "backward", trace = TRUE)
summary(backmod)
```

Does this model differ from the forward model? If so, how?
// it does not include the working day variable, being a backward stepwise, it is coverded by weekend and holiday, because it will be a working day if it is neither weekend or holiday.

```{r}
bike2 = bike2 %>% mutate(yr = as.integer(yr)-1)
```

```{r}
allmod1 = lm(count ~., bike2)
summary(allmod)

emptymod2 = lm(count ~1, bike2)
summary(emptymod)
```

```{r}
forwardmod1 = stepAIC(emptymod2, direction = "forward", scope=list(upper=allmod1,lower=emptymod2),
                      trace = TRUE)
summary(forwardmod1)
```

Did the model change?
// R-squared and p value remained the same. Same variables included in the model.


### Brian Andrino








