---
output:
  word_document: default
  html_document: default
---
```{r}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(nnet)
library(caret)
```

```{r}
parole = read_csv("parole.csv")
```

```{r}
parole = parole %>% mutate(male = as_factor(as.character(male))) %>%
mutate(male = fct_recode(male,"male" = "1","female" = "0")) 
```

```{r}
parole = parole %>% mutate(race = as_factor(as.character(race))) %>%
mutate(race = fct_recode(race,"white" = "1","otherwise" = "2")) 
```

```{r}
parole = parole %>% mutate(state = as_factor(as.character(state))) %>%
mutate(state = fct_recode(state,"other" = "1","Kentucky" = "2", "Louisiana" = "3", "Virginia" = "4")) 
```

```{r}
parole = parole %>% mutate(crime = as_factor(as.character(crime))) %>%
mutate(crime = fct_recode(crime,"other" = "1","larceny" = "2", "drug-related" = "3", "driving-related" = "4")) 
```

```{r}
parole = parole %>% mutate(multiple.offenses = as_factor(as.character(multiple.offenses))) %>%
mutate(multiple.offenses = fct_recode(multiple.offenses,"multiple.offenses" = "1","otherwise" = "0")) 
```

```{r}
parole = parole %>% mutate(violator = as_factor(as.character(violator))) %>%
mutate(violator = fct_recode(violator,"violation" = "1","no-violation" = "0")) 
```

```{r}
set.seed(12345) 
train.rows = createDataPartition(y = parole$violator, p=0.7, list = FALSE)
train = parole[train.rows,] 
test = parole[-train.rows,]
```

```{r}
fitControl = trainControl(method = "cv", 
                           number = 10)

nnetGrid <-  expand.grid(size = 12, decay = 0.1)

set.seed(1234)
nnetBasic = train(violator ~ ., 
                 parole,
                 method = "nnet",
                 tuneGrid = nnetGrid,
                 trControl = fitControl,
                 verbose = FALSE,
                 trace = FALSE)
```

```{r}
nnetBasic
```

```{r}
predNetBasic = predict(nnetBasic, train)
```

```{r}
confusionMatrix(predNetBasic, train$violator)
```

Comment on the model quality.
// This seems to be a good model, the accuracy level is high and is consistent with other model that we have developed before.

```{r}
fitControl2 = trainControl(method = "cv", 
                           number = 10)

nnetGrid2 =  expand.grid(size = seq(from = 1, to = 12, by = 1), 
                        decay = seq(from = 0.1, to = 0.5, by = 0.1))
set.seed(1234)
nnetFit = train(violator ~ ., 
                 parole,
                 method = "nnet",
                 trControl = fitControl2,
                 tuneGrid = nnetGrid2,
                 verbose = FALSE,
                trace = FALSE)
```

```{r}
nnetFit
```

```{r}
predNet = predict(nnetFit, train)
```

```{r}
confusionMatrix(predNet, train$violator)
```

```{r}
predNetBasic_test = predict(nnetBasic, newdata = test)
```

```{r}
confusionMatrix(predNetBasic_test, test$violator)
```

```{r}
predNet_test = predict(nnetFit, newdata = test)
```

```{r}
confusionMatrix(predNet_test, test$violator)
```

Comment on whether there appears to be overfitting in one or both of your models from Tasks 2 and 4.

The accuracy levels from task 2 are 0.9302 for the training set and 0.9158, the accuracy level from task 4 are 0.9049 for the training set and 0.9061 for the testing set. In my opinion neither of the models look too overfitting, however, task 4 shows a greater difference in accuracy between training and testing sets.

